---
title: "P8105 Homework 3"
author: Chenxin Zhang
date: 10/10/2020
output: html_document
---

```{r setup, message = FALSE, echo = FALSE}
library(p8105.datasets)
library(tidyverse)
library(readxl)
library(dplyr)
library(rnoaa)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Probalem 1
```{r message = FALSE, echo = FALSE}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r message = FALSE, echo = FALSE}
# count how many aisle 
# arrange the number of items of aisles  by decreasing
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r message = FALSE, echo = FALSE}
# limiting this to aisles with more than 10000 items
# theme()rearrange x-axis
# how ggplot put things in order: reorder facter aisle by n
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r message = FALSE, echo = FALSE}
#group product name by aisles at first then  count
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```

Apples vs ice cream..

```{r message = FALSE, echo = FALSE}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

## Probalem 2
#### 2.1 Load, tidy, and otherwise wrangle the data
```{r message = FALSE, echo = FALSE}
#Load, tidy, and otherwise wrangle the data
#include a weekday vs weekend variable
#describe the resulting dataset 
accel_df =
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "ac_counts") %>% 
  mutate(
    week = as.factor(week),
    minute = as.integer(minute),
    day = as.factor(day),
    day_id = as.factor(day_id) ) %>% 
  
  mutate(
    weekday_weekend = case_when(
      day %in% c("Monday","Tuesday","Wednesday","Turthday","Friday") ~ "weekday",
      day %in% c("Saturday", "Sunday") ~ "weekend"), 
    weekday_weekend = factor(weekday_weekend)) %>% 
  
  mutate(
    day = forcats::fct_relevel(day,"Monday","Tuesday","Wednesday","Thursday","Friday", "Saturday","Sunday")) 
accel_df
```

* This dataset includes `r nrow(accel_df)` rows and `r ncol(accel_df)`. It is consist of variabels：`r names(accel_df)`.

#### 2.2 the total activity over the day
```{r message = FALSE, echo = FALSE}
#the total activity over the day.
accel_df2 = accel_df %>% 
  group_by(week, day) %>% 
  summarize(total_counts = sum(ac_counts)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_counts) %>% 
  mutate(total_week = sum(Monday:Sunday)) %>% 
  knitr::kable(digits = 2)
accel_df2
```
*The table shows that the men does less activities on weekend than on weekdays.

#### 2.3 single-panel plot
```{r message = FALSE, echo = FALSE}
#a single-panel plot that shows the 24-hour activity time courses for each day 
#use color to indicate day of the week
#Describe in words any patterns or conclusions 
accel_df3 = accel_df %>% 
  ggplot(aes(x = minute, y = ac_counts, color = day)) +
  geom_smooth(se = F) +
  labs(
    title = "Total activities in a day over 35 days",
    caption = "Data from accelerometer") +
   viridis::scale_color_viridis(discrete = T,
                               name = "Day") +
  scale_x_continuous(name = "Time(h)",
                    breaks = c(120, 240, 360, 480, 600, 720, 840, 960, 1080, 1200, 1320, 1440),
                    labels = c("2h", "4h", "6h", "8h", "10h", "12h","14h", "16h", "18h","20h", "22h", "24h")) +
  scale_y_continuous(name = "Activities in minutes",trans = "sqrt")
accel_df3
```

* As is shown in the graph, the men usually does activity frequently from 8am to 9pm in weekdays. During weekend, he has less activity time on Saturday, while much activity time in Sunday morning. 

## Problem 3
#### 3.1 Data cleaning 
```{r message = FALSE, echo = FALSE}
data("ny_noaa")
```

```{r data cleaning, message = FALSE, echo = FALSE}
#Create separate variables for year, month, and day
#temperature, precipitation, and snowfall are given in reasonable units
ny_noaa_df = ny_noaa %>% 
  mutate_at(vars(date), as.factor) %>% 
  mutate_at(vars(tmax, tmin, prcp, snow), as.numeric) %>%
  separate(date, into = c("year", "month", "day")) %>% 
  mutate_at(vars(year, month, day), as.factor) %>% 
  mutate(
    prcp = prcp/10,
    tmax = tmax/10,
    tmin = tmin/10,
    snow = case_when(
      snow <0 ~0,
      snow >= 0 ~snow))
skimr::skim_without_charts(ny_noaa_df)
```

```{r snowfall, message = FALSE, echo = FALSE}
snow_df = ny_noaa_df %>% 
  count(snow, na.rm = T) %>% 
  mutate(rank = min_rank(desc(n))) 
snow_df
```

* The dataset is consist of `r nrow(ny_noaa_df)` rows and `r ncol(ny_noaa_df)` columns. It contains variables: `r names(ny_noaa_df)`. For snowfall, the most commonly observed value is 0. As is shown in the snow_df, there are 2,008,509 out of 2,595,176 days without snow, which rank 1.


#### 3.2 Average max temperature in January and in July 

```{r Jan_avg_temp, message = FALSE, echo = FALSE}
Jan_avg_temp = ny_noaa_df %>% 
  filter(month == "01") %>% 
  group_by(year, month, id) %>% 
  summarise(avg_tmax = mean(tmax,na.rm = T)) %>% 
  drop_na() %>% 
ggplot(aes(x = year, y = avg_tmax, color = id)) +
  geom_point(alpha = 0.3, size = 0.1) +
  geom_path(aes(group = id), alpha = 0.3) +
  theme(
    plot.title = element_text(lineheight = 4, color = "black"),
    legend.position = 'none',
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "The average max temperature in January in each station across years",
    x = "Year",
    y = "Average max temperature")
```
*January outliers
```{r january outliers, message = FALSE, echo = FALSE, collapse = TRUE} 
ny_noaa_df %>% 
  filter(month == "01") %>% 
  group_by(year, month, id) %>% 
  summarise(avg_tmax = mean(tmax,na.rm = T)) %>% 
  drop_na() %>% 
  filter(avg_tmax > 10 | avg_tmax < -10) %>% 
  knitr::kable(digits = 1)
```

```{r message = FALSE, echo = FALSE}
jul_avg_temp = ny_noaa_df %>% 
  filter(month == "07") %>% 
  group_by(year, month, id) %>% 
  summarise(avg_tmax = mean(tmax,na.rm = T)) %>% 
  drop_na() %>% 
ggplot(aes(x = year, y = avg_tmax, color = id)) +
  geom_point(alpha = 0.3, size = 0.1) +
  geom_path(aes(group = id), alpha = 0.3) +
  theme(
    plot.title = element_text(lineheight = 4, color = "black"),
    legend.position = 'none',
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "The average max temperature in July in each station across years",
    x = "Year",
    y = "Average max temperature")
```

*July outliers
```{r july outliers, message = FALSE, echo = FALSE, collapse = TRUE} 
ny_noaa_df %>% 
  filter(month == "07") %>% 
  group_by(year, month, id) %>% 
  summarise(avg_tmax = mean(tmax,na.rm = T)) %>% 
  drop_na() %>% 
  filter(avg_tmax > 33 | avg_tmax < 20) %>% 
  knitr::kable(digits = 1)
```

```{r}
Jan_avg_temp/jul_avg_temp
```

* "The average max temperature in July in each station across years" shows that it is relatively low temperature in 1994,2004, and relatively high temperature in 1990,1995,1998,2002,2006. "The average max temperature in July in each station across years" shows that it is relatively low temperature in 1988,1992,2000 and relatively high temperature in 1983,1999,2010. According to the graph, we can easily see that average max temperature in both January and July over year are fluctuated. The temperature fluctuate between -13.4 to 10.2, while the temperature fluctuate between 14.0 to 33.6.

#### 3.3 
*plot showing tmax vs tmin for the full dataset 
*plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
```{r message = FALSE, echo = FALSE, collapse = TRUE}
temp_df = ny_noaa_df %>% 
  drop_na() %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "tmax_tmin",
    values_to = "temperature") %>% 
  ggplot(aes(x = year, y = temperature)) +
  geom_boxplot(aes(color = tmax_tmin), alpha = 0.4, outlier.size = 0.1)+
  theme(
    legend.position = 'right',
    plot.title = element_text(lineheight = 4, color = "black", size = 20),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 10)) +
  labs(title = "The average max temperature in July in each station across years",
    x = "Year",
    y = "tmax vs tmin for the full dataset")
```

```{r message = FALSE, echo = FALSE, collapse = TRUE}
snowfall_df = ny_noaa_df %>% 
  filter(snow < 100 & snow > 0) %>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = snow)) + 
  geom_violin(color = "black", fill = "green", alpha = .5)+
  theme(
    plot.title = element_text(lineheight = 4, color = "black", size = 20),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    title = "The distribution of snowfall values over year",
    x = "Year",
    y = "Snowfall (mm)")
```
*two-panel plot
```{r}
(temp_df / snowfall_df) + plot_layout(widths = 10, heights = 20)
```

* The temp_df shows that the maximum temperatures over years have a larger range between the Q1 to Q3 than the minimum temperature, according to the width of the box. However, the minimum temperatures over years have larger range from Q1 to Q4, as they have more outliers.
* The snowfall_df shows the values of snowfall concentrate on 0-25mm, 50mm, 75mm.  The fluctuation of snowfall over years are stable. 


